---
output:
  html_document: 
    # highlight: pygments
    toc: yes
    toc_float: yes
    toc_depth: 4
    number_sections: yes
---

## Code

**"Rutil" is my collection of utility and can be viewed online at https://github.com/shouldsee/Rutil/**

This code appendix is also submitted as a html separately.

### Fitting the beta distribution to constraints

`r fname<-'fitting_beta.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

### Systematic helpers for inference on the [0,1] interval

`r fname<-'bayes_infer.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

### Bayesian network inference

`r fname<-'dag.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

### Construct posterior accroding to the graphical model

`r fname<-'net_post.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

#### Graphs

`r fname<-'graph_0edge.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

`r fname<-'graph_1edge.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

`r fname<-'graph_2edge.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

### Comparison of algortihms

#### Homemade xgboost-based regressional predictor (Gxgb)

`r fname<-'xgb.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

#### Calculation of QC-metrics 

`r fname<-'dream5.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

#### Routines for commonly used algorithms

`r fname<-'routine.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

#### Bootstrapping on sub-networks

`r fname<-'subnet.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

##### Working script for GNW dataset

`r fname<-'assignment-subnet.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

##### Working script for DREAM5 dataset, network 3

`r fname<-'ecoli-subnet.R'`

```{r,code=readLines(fname),eval = F,echo = T}
### CODE
```

